---
title: "Sequence analysis"
output: html_document
date: '2022-09-23'
---

# Reference 

https://greendaygh.github.io/kribbr2022/


# Set environment  

1. Installation R/Rstudio on your computer
2. Create a directory "smmo" for the analysis
3. Create a project "smmo"
4. Install CRAN packages (tidyverse, ...)
    - by seeing (https://greendaygh.github.io/kribbr2022/rrstudio-basics.html), do 1, 2, 3
5. Install bioconductor packages (Biostrings, GenomicRanges, ShortReads, ...) 
    - https://greendaygh.github.io/kribbr2022/bioconductor.html you can do 4
6. Create a "data" directory for reference genomic sequence of M. capsulatus Bath 
    - download .fasta, .gb (genbank file of the strain if possible) files from NCBI
7. Copy fastq files to the data directory 


# Sample description 



# fastq file loading


- 일반적으로 ShortRead 패키지 사용, 본 교육에서는 bioconductor의 `Rfastp` 패지키 사용. 
- Install Rfastp from bioconductor if necessary. 

```{r, eval=F}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Rfastp")
```


- fastq 파일 다운로드 as an example
- `rfastp` file load / filtering


```{r, eval=F}
library(Rfastp)

fqfiles <- dir(path = "fastq", pattern = "*.fastq.gz")

for(i in 1:length(fqfiles)){
  cat(i, "\n");flush.console()
  json_report <- rfastp(read1 = file.path("fastq", fqfiles[i]), 
                      outputFastq = file.path("outfastq", fqfiles[i]))
}
#qcSummary(json_report)

``
`


# prepare a reference sequence  

```{r, eval=F}
library(Biostrings)

ftnseq <- readDNAStringSet("data/pACBB-Standard FTN sensor.fa")
petseq <- readDNAStringSet("data/pACBB-Standard PET sensor.fa")

```

# create index for mapping

- align 전에 index를 생성해야 함 
- Rsubread 패키지 사용

```{r, eval=F}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Rsubread")
```

- becareful if you're working with a big sequence data
- build index

```{r, eval=F}
library(Rsubread)

buildindex("mm10_smallchars", "myref.fa")
```

# mapping


```{r, eval=F}
library(Rsubread)

alignstat <- align("data/mm10_smallchars",
                   readfile1 = "data/ENCFF332KDA_rfastp_R1.fastq.gz",
                   nthreads = 10)


alignstat

```

# sorting


```{r, eval=F}

library(Rsamtools)

sortBam(file = "data/ENCFF332KDA_rfastp_R1.fastq.gz.subread.BAM", 
        "data/Sorted_ENCFF332KDA_rfastp_R1.fastq.gz.subread.BAM")

indexBam("data/Sorted_ENCFF332KDA_rfastp_R1.fastq.gz.subread.BAM.bam")

```



# pileup

```{r, eval=F}

sbp <- ScanBamParam(which=GRanges("pPROLar-pac", IRanges(1, end_coord)))
pparam <- PileupParam(max_depth=10000,
                      distinguish_nucleotides = TRUE,
                      include_insertions=TRUE,
                      distinguish_strands = FALSE, 
                      min_nucleotide_depth=0,
                      min_mapq=0,
                      min_base_quality=0)
res <- pileup(filename, scanBamParam=sbp, pileupParam=pparam)
#head(res)
#res_sel <- res %>% filter(pos>=138 & pos<=2675) ## PenG 
## convert table to data.frame
res_table <- as.data.frame.matrix(table(res$pos, res$nucleotide))
#plot(rowSums(res_table[,1:4]), type="h")


```

