---
title: "Sequence analysis"
output: html_document
date: '2022-09-23'
---

# Reference 

https://greendaygh.github.io/kribbr2022/


# Set environment  

1. Installation R/Rstudio on your computer
2. Create a directory "smmo" for the analysis
3. Create a project "smmo"
4. Install CRAN packages (tidyverse, ...)
    - by seeing (https://greendaygh.github.io/kribbr2022/rrstudio-basics.html), do 1, 2, 3
5. Install bioconductor packages (Biostrings, GenomicRanges, ShortReads, ...) 
    - https://greendaygh.github.io/kribbr2022/bioconductor.html you can do 4
6. Create a "data" directory for reference genomic sequence of M. capsulatus Bath 
    - download .fasta, .gb (genbank file of the strain if possible) files from NCBI
7. Copy fastq files to the data directory 


# Sample description 

- NGS (library) Samples: 3 fastq files 
- 5k~6kbp, 5~6 genes, SMMO in genome
- Sent cell pellets, extracted genomic DNA and sequencing
- sample1: high fluorescence ?? reaction time 
- sample2: high fluorescence ?? reaction time 
- sample3: low fluoresscence 



# fastq file loading


일반적으로 ShortRead 패키지 사용, 본 교육에서는 bioconductor의 `Rfastp` 패지키 사용. Install from bioconductor if necessary. 

```{r, eval=F}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Rfastp")
```


파일을 다운로드하고 `rfastp` 파일로 읽음. 


```{r, eval=F}
library(Rfastp)

fq <- "https://www.encodeproject.org/files/ENCFF332KDA/@@download/ENCFF332KDA.fastq.gz"
download.file(fq, "ENCFF332KDA.fastq.gz")


json_report <- rfastp(read1 = "ENCFF332KDA.fastq.gz", outputFastq = "ENCFF332KDA_rfastp")
qcSummary(json_report)

```


# Alignment 

